---
title: "EFForTS-ABM landmarket summary"
author: "Jan Salecker"
date: "2019 M04 3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This repository will summarize the analysis on newly introduced landmarket feature within the EFForTS-ABM.

We need to find a good story.

Interesting interactions are:
landmarkets affect the distribution of wealth and land-uses.
Heterogeneity and learning affect landmarkets.
Prices and Landscapes affect landmarkets.

As model input we could use:
* Different landscapes
* Different Price scenarios (smoothing)

As model output we could look at:
* Distribution of land-uses (landscape metrics)
* Distribution of wealth
* Aggregated metrics (consolidation, immigration)

Maybe we could relate the distribution of wealth to other metrics, e.g. carbon/biodiversity/inputs


## Hypotheses:

* H1: Landmarkets lead to consolidation
* H1.1: Large smallholders are more likely to buy land 


## Analyses:

* Start with simple model: no heterogeneity, no learning, historical prices

* H1: Landmarkets lead to consolidation
** Measure household sizes and landmarket events over time

* H1.1: smallholders with large area are more likely to buy land, smallholders with small area are more likely to sell land
** Create output when landmarkets happen about the propertie of all households in the buyer pool
** Measure: h_wealth, h_area, h_inefficiencies, h_immigrant?
** Create a plot about properties of selling households and buying households over time


```{r echo=FALSE}
library(DiagrammeR)
DiagrammeR::grViz("digraph {

graph [layout = dot, rankdir = LR]

# define the global styles of the nodes. We can override these in box if we wish
node [shape = rectangle, style = filled, fillcolor = Linen]

input1 [label = 'Landscapes', shape = folder, fillcolor = Beige]
input2 [label = 'Prices', shape = folder, fillcolor = Beige]
ABM [label =  'EFForTS-ABM', fillcolor = gold]
output1 [label = 'Landmarket effects']
output2 [label= 'hh properties']

# edge definitions with the node IDs
{input1 input2}  -> ABM -> {output1 output2}
}")

```



